---
title: "Classifying musical genre using Spotify audio features"
date: 2020-07-02
output: html_document
editor_options:
  chunk_output_type: console
---

This analysis attempts to classify songs into their correct musical genre using
audio features. It is inspired by the original analysis by Kaylin Pavlik
([\@kaylinquest][kaylinquest]) in her 2019 blog post [Understanding +
classifying genres using Spotify audio features][blog-post].

[kaylinquest]: https://twitter.com/kaylinquest
[blog-post]: https://www.kaylinpavlik.com/classifying-songs-genres/
```{r}
library(rpart) #EDJ
```

```{r import}
#spotify <- read.csv("/Users/login/Dropbox/spotify/spotify.csv", stringsAsFactors = FALSE)
# correct file path and drop string args (default is TRUE in R 4.0)
spotify <- read.csv("data/spotify.csv") #EDJ
dim(spotify)
head(spotify)
#table(spotify[, 1])
table(spotify$genre) #EDJ
spotify <- spotify[, 1:13]
```

Split the data into training and testing sets.

```{r split}
#where did numTrainingSamples come from?
#numTrainingSamples <- 24624 #EDJ
numTrainingSamples <- .75*nrow(spotify)
#trainingSet <- sample(seq_len(nrow(spotify)), size = numTrainingSamples)
trainingSet <- sample(nrow(spotify), size = numTrainingSamples) #EDJ
spotifyTraining <- spotify[trainingSet, ]
spotifyTesting <- spotify[-trainingSet, ]
```

Build classification model with decision tree from the rpart package.

```{r classification}
model <- rpart(genre ~ ., data = spotifyTraining)
plot(model)
text(model)
```

Calculate prediction accuracy of the model on the training and testing sets.

```{r accuracy}
predictTraining <- predict(model, type = "class")
(accuracyTraining <- mean(spotifyTraining[, 1] == predictTraining))

predictTesting <- predict(model, newdata = spotifyTesting[, -1], type = "class")
(accuracyTesting <- mean(spotifyTesting[, 1] == predictTesting))
```

Evaluate prediction performance using a confusion matrix.

```{r confusion-matrix}
table(predicted = predictTesting, observed = spotifyTesting[, 1])
```

How does the model compare to random guessing?

```{r random-guess}
predictRandom <- sample(unique(spotifyTesting[, 1]),
                        size = nrow(spotifyTesting),
                        replace = TRUE,
                        prob = table(spotifyTesting[, 1]))
#(accuracyRandom <- mean(spotifyTesting[, 1] == predictGuess))
(accuracyRandom <- mean(spotifyTesting[, 1] == predictRandom)) #EDJ
```
